Based on polling as of 27 October 2024, Harris expects to win on xxx%, leads Trump by xx% in 2024 US election

### 2024-10-30 paper/paper.qmd
Add discussion point

---
title: "Based on polling as of 27 October 2024, Harris expects to win on 49.2%, leads Trump by 2% in 2024 US election"
subtitle: "Weighted Linear Regression Analysis of Aggregated Poll Data: Utilizing National Polls indicator, Population, and Pollster as Predictors."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(arrow)
library(dplyr)
```

```{r}
#| include: false
#| warning: false
#| message: false
harris_data <- read_parquet("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Harris.parquet")
trump_data <- read_parquet("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Trump.parquet")

@@

```{r}
#| include: false
#| warning: false
#| message: false
harris_unweighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_harris_unweighted.rds")
harris_weighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_harris_weighted.rds")
trump_unweighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_trump_unweighted.rds")
trump_weighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_trump_weighted.rds")

@@

@@

```{r}
#| include: false
#| warning: false
#| message: false
#### Calculate Weights ####
# Calculate weights for Harris data
harris_data <- harris_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)
```

@@

1.  **Use of Weighted Model**: The weighted model is essential for
    accurately reflecting voter sentiment. By incorporating weights, we
    adjust for the reliability of the polling data, ensuring that more
    credible polls have a greater influence on the estimates.

@@

### Weighted Least Squares Estimation

@@

# Load necessary libraries
library(tidyverse)
library(ggplot2)

harris_data <- harris_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")

recent_date_cutoff <- as.Date("2024-08-01")  # Set cutoff for recent data

harris_filtered <- harris_data |>
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Harris")
trump_filtered <- trump_data |>
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Trump")

combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-national-support-trends
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States
harris_data<-harris_data|>mutate(candidate = "Harris")
trump_data<-trump_data |>mutate(candidate= "Trump")
combined_data_total <- bind_rows(harris_data, trump_data)
combined_data_total|>filter(national_poll==1)|>ggplot(aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Country",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-27"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )

new_data_trump <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )
```

```{r}
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(model_harris_unweighted, newdata = new_data_harris),
    predicted_weighted = predict(model_harris_weighted, newdata = new_data_harris)
  )

# Predictions for Trump
new_data_trump <- new_data_trump |>
  mutate(
    predicted_unweighted = predict(model_trump_unweighted, newdata = new_data_trump),
    predicted_weighted = predict(model_trump_weighted, newdata = new_data_trump)
  )

# Plot the predictions
library(dplyr)
library(ggplot2)

# Define the cutoff date for recent data
recent_date_cutoff <- as.Date("2024-05-01")

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions_filtered <- combined_predictions %>%
  filter(end_date >= recent_date_cutoff)

# Determine the maximum end date for setting the x-axis limit
max_end_date <- max(combined_predictions_filtered$end_date)

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = FALSE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = FALSE, size = 1) +  # Weighted model with smoothing
  labs(title = "Predicted Support for Kamala Harris and Donald Trump \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

```
The predicted support for Kamala Harris and Donald Trump, as derived from multiple linear regression models, is displayed in the graph above. The analysis spans from August 1, 2024, to October 27, 2024, and highlights both unweighted and weighted predictions for each candidate.

For Kamala Harris, the unweighted model predicts her support to remain relatively stable, maintaining an approximate support level of 49%. This trend is visually represented by the blue dashed line, which indicates minimal fluctuations over time. Conversely, the weighted model shows a more pronounced trend, demonstrating a gradual increase in predicted support, suggesting a potential shift in voter sentiment towards Harris as the election date approaches.

In contrast, Donald Trump's predicted support, represented by the red lines, exhibits a downward trend in both unweighted and weighted models. The unweighted model indicates that his support is expected to hover around 48%, while the weighted model predicts a slight decline to approximately 46%. This trend may reflect challenges in maintaining voter enthusiasm in the face of changing public sentiment.

At the end of the polling period, we predict that Kamala Harris leads Donald Trump by 2% in the unweighted model (48% vs. 46%) and by about 2% in the weighted model (49.1% vs. 47.3%).

@@

## First discussion point {#sec-first-point}

@@

Recency weights are calculated based on the number of days elapsed since
the poll ended, utilizing an exponential decay function to give greater
weight to more recent polls.

@@

## Model validation: K-Fold Cross-Validation

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
# Load necessary libraries
library(boot)    # For bootstrapping
library(caret)   # For cross-validation

# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)
# Train the model
model_formula <- pct ~ national_poll + pollster + population
model_harris_cv <- train(model_formula, data = harris_data, method = "lm", trControl = train_control)
# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)
# Train the model
model_harris_cv <- train(model_formula, data = harris_data, method = "lm", trControl = train_control)
# Print the model results
print(model_harris_cv)

model_trump_cv <- train(model_formula, data = trump_data, method = "lm", trControl = train_control)
# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)
# Train the model
model_trump_cv <- train(model_formula, data = trump_data, method = "lm", trControl = train_control)
# Print the model results
print(model_trump_cv)
```

We use a 10-fold cross-validation on two linear regression models—one
for Harris and one for Trump. The models use three predictors:
national_poll, pollster, and population. The output provides key
metrics, which breaks down here:

RMSE (Root Mean Square Error): Measures the average magnitude of
prediction errors (lower is better).

Harris model: RMSE of 3.47, indicating an average prediction error of
around 3.47 percentage points.\
Trump model: RMSE of 4.39, showing a slightly higher prediction error on
average.\

R-squared: Represents the proportion of the variance in the response
variable explained by the model (higher is better).\
Harris model: R-squared of 0.346, meaning the model explains about 34.6%
of the variance in Harris's polling data.\
Trump model: R-squared of 0.346 as well, indicating similar explanatory
power for Trump's polling data.

MAE (Mean Absolute Error): Shows the average absolute difference between
observed and predicted values (lower is better).\
Harris model: MAE of 2.50, meaning that, on average, the predictions are
off by 2.5 percentage points.\
Trump model: MAE of 3.13, indicating slightly less precise predictions
compared to the Harris model.\

Interpretation Summary Predictive Accuracy: The Harris model has
slightly better predictive accuracy than the Trump model, as reflected
by its lower RMSE and MAE values.\
Model Fit: Both models explain roughly 34.6% of the variance in their
respective datasets. This suggests that other factors not included in
the model may play a significant role in explaining the remaining
variance.\
This summary indicates the models are moderately predictive, with room
for improvement in accuracy and fit, potentially by adding more
predictors or adjusting model specifications.\

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\-Support for Harris"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

# Residual plot for unweighted model
plot(harris_unweighted_model, which = 1, main = "Residual Plot of Unweighted Linear Regression Model")

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(harris_unweighted_model), main = "Q-Q Plot of Unweighted Linear Regression Model Residuals")
qqline(residuals(harris_unweighted_model), col = "red")

# Residual plot for weighted model
plot(harris_weighted_model, which = 1, main = "Residual Plot of Weighted Linear Regression Model")

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(harris_weighted_model), main = "Q-Q Plot of Weighted Linear Regression Model Residuals")
qqline(residuals(harris_weighted_model), col = "red")

@@

Generally, we use Residual vs Fitted plot and Q-Q Plot diagnostic our
model. Residual vs Fitted plot aare Residuals (differences between
observed and predicted values) plotted against fitted values. Ideally,
these residuals should be randomly scattered around the zero line to
indicate that the model does not have systematic errors. The Q-Q plot
for the unweighted model shows how the residuals align with a
theoretical normal distribution. Ideally, residuals should follow a
straight line in this plot if they are normally distributed, which is an
assumption of linear regression.

@fig-stanareyouokay-1 is a residual plot of un-weighted model for Harris
support. It showsthe residuals are generally spread around zero, with no
clear pattern. This suggests the model is relatively well-specified.
However, there is a slight curvature, indicating potential non-linearity
that the model may not fully capture. A few notable outliers with larger
residuals might be influencing the model, indicating that some data
points have more significant prediction errors.

@fig-stanareyouokay-2 is a Q-Q plot of un-weighted model plot for Harris
support. It shows most residuals fall along the line, especially in the
middle range. This suggests that our model satisfy the normality
assumption. However, some points at the tails deviate, indicating
potential outliers or non-normality in the extreme residual values. This
slight deviation at the ends suggests the model might have some issues
with extreme predictions but performs reasonably well overall.

@fig-stanareyouokay-3 is a residual plot of weighted model for Harris
support.It shows residuals are again plotted against fitted values.
Similar to the unweighted model, the residuals are mostly centered
around zero, indicating that the weighted model captures the general
trend without significant systematic bias. The curvature is slightly
reduced compared to the un-weighted model, suggesting that weighting has
helped in addressing some of the non-linearity observed in the
un-weighted model. However, some residuals are still notably large,
which may indicate outliers that influence the model despite the
weighting scheme. This suggests that while the weighted model performs
better in terms of capturing non-linearity, further refinement might
still be beneficial.

@fig-stanareyouokay-4 is a Q-Q Plot of weighted model plot for Harris
support.The residuals generally align with the theoretical normal
distribution line, particularly in the central range, indicating that
the residuals of the weighted model are close to normal. Similar to the
unweighted model, there are deviations at the tails, though they appear
less pronounced. This suggests that the weighting scheme has slightly
improved the distribution of residuals, making the model’s predictions
more robust. However, some extreme values remain, which could still
affect model

In summary, both models show a reasonably good fit, with the weighted
model offering slight improvements in handling non-linearity and extreme
values. However, both models exhibit minor deviations from normality and
a few notable outliers, which may warrant further model adjustments for
improved prediction accuracy.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-dia
#| fig-cap: "Diagnostics of model using residual vs fitted plot and norm Q-Q plot\n-Support for Trump"
#| fig-subcap: ["Residual Plot for Unweighted Model", "Q-Q Plot for Unweighted Model", "Residual Plot for Weighted Model", "Q-Q Plot for Weighted Model"]
#| layout-ncol: 2

# Residual plot for unweighted model
plot(trump_unweighted_model, which = 1, main = "Residual Plot of Unweighted Linear Regression Model")

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(trump_unweighted_model), main = "Q-Q Plot of Unweighted Linear Regression Model Residuals")
qqline(residuals(trump_unweighted_model), col = "red")

# Residual plot for weighted model
plot(trump_weighted_model, which = 1, main = "Residual Plot of Weighted Linear Regression Model")

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(trump_weighted_model), main = "Q-Q Plot of Weighted Linear Regression Model Residuals")
qqline(residuals(trump_weighted_model), col = "red")

@fig-dia-1 shows the residuals plotted against the fitted values for the
unweighted model. It shows that the residuals are generally spread
around zero with no clear pattern, suggesting that the model is
relatively well-specified. However, there is a slight curvature,
indicating possible non-linearity that the model may not fully capture.
Some notable outliers with larger residuals suggest that certain data
points have significant prediction errors, potentially influencing the
model.

@fig-dia-2 shows the Q-Q plot for the unweighted model, showing that
most residuals fall along the line, especially in the middle range,
suggesting that the model satisfies the normality assumption. However,
some points at the tails deviate, indicating potential outliers or
non-normality in extreme residual values. This deviation at the ends
suggests the model may face issues with extreme predictions, though it
performs reasonably well overall.

@fig-dia3 shows residuals plotted against fitted values. Similar to the
unweighted model, the residuals are centered around zero, indicating
that the weighted model captures the overall trend without significant
systematic bias. The slight curvature seen in the unweighted model is
reduced here, suggesting that weighting has addressed some of the
non-linearity. However, some large residuals remain, which could
indicate outliers that affect the model even with the weighting scheme.
This suggests that while the weighted model has improved in handling
non-linearity, further refinements could enhance accuracy.

@fig-dia-4 shows the Q-Q plot for the weighted model compares residuals
with a theoretical normal distribution. Here, the residuals generally
align with the theoretical line, especially in the central range,
indicating that the residuals of the weighted model are close to normal.
Similar to the unweighted model, some deviations occur at the tails,
though they appear less pronounced, suggesting that the weighting scheme
has slightly improved the normality of residuals. Nonetheless, some
extreme values persist, which may impact model robustness in cases of
outliers.

Summary Both models exhibit a reasonably good fit, with the weighted
model offering slight improvements in managing non-linearity and extreme
values. Despite this, both models show minor deviations from normality
and some notable outliers, suggesting that further model adjustments may
be beneficial for improved prediction accuracy.

>>>>

---
title: "Based on polling as of 27 October 2024, Harris expects to win on xxx%, leads Trump by xx% in 2024 US election"
subtitle: "Weighted Linear Regression Analysis of Aggregated Poll Data: Utilizing National Polls indicator, Population, and Pollster as Predictors."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
```

@@

@@

```{r}
model_harris_unweighted <- lm(pct ~ national_poll + pollster + population, data = harris_data)
```

@@

```{r}
model_harris_weighted <- lm(pct ~ national_poll + pollster + population, data = harris_data, weights = combined_weight)
```

@@

1. **Weighted Least Squares Estimation**

The weighted model is essential for accurately reflecting voter sentiment. 
By incorporating weights, we adjust for the reliability of the polling data, 
ensuring that more credible polls have a greater influence on the estimates.

@@

@@

harris_data <- harris_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

harris_data <- harris_data %>% mutate(candidate = "Harris")
trump_data <- trump_data %>% mutate(candidate = "Trump")
combined_data_total <- bind_rows(harris_data, trump_data)

selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
recent_date_cutoff <- as.Date("2024-08-01")  # Set cutoff for recent data

harris_filtered <- harris_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

trump_filtered <- trump_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-national-support-trends
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)

combined_data_total|>filter(national_poll==1)|>ggplot(aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Country",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-27"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))

@@

@@

## Key Findings and Real-World Implications {#sec-key-findings}
Our model reveals a portrait of a nation deeply divided, with Harris holding a narrow 49% edge over Trump’s 47% in national support. However, as recent elections have shown, the national popular vote does not always translate into an Electoral College victory. The U.S. electoral system, which awards each state’s electoral votes to the candidate with the majority of votes in that state, means that winning the popular vote nationally might still leave Harris short of the presidency. This structural quirk played a defining role in the 2016 election, where Hillary Clinton’s popular vote win failed to deliver the electoral majority, paving Trump’s path to the White House. Our findings underscore this enduring tension: while Harris may have a slight national advantage, the true battle will be fought state by state, in a handful of battlegrounds that could flip the election either way.

In states like Wisconsin and Arizona, our model highlights just how close the race remains. Wisconsin, for instance, shows both candidates locked in a near tie around 48% support—a statistical dead heat that brings the state’s significance into sharp relief. Arizona, meanwhile, leans modestly toward Trump, a reflection of the unique demographic and political nuances shaping each battleground. These state-level insights illuminate a critical truth: while national polls offer a snapshot of overall sentiment, they risk obscuring the specific, local dynamics that will ultimately decide the outcome. The stakes are high; these are the states that could swing, the margins that will be watched closely on election night, as even slight changes in turnout or last-minute shifts in opinion could tip the balance.

Yet, as with all models, limitations persist. While our weighting system, with its emphasis on recent data, aims to capture a timely snapshot of voter sentiment, it may miss sudden changes sparked by campaign events or unexpected shifts in public discourse. The variability in our state projections—ranging from a low estimate of 363 electoral votes for Harris to a high of 471—reflects the inherent uncertainty in polling-based models, particularly given the constraints of our data sources. This wide range suggests that while data-driven insights can illuminate trends, they cannot fully account for the unpredictable nature of electoral outcomes. Moving forward, incorporating real-time sentiment from social media or demographic-specific insights could offer a more responsive understanding of voter behavior. Ultimately, these findings highlight the intricate dance between popular sentiment and the electoral mechanics of American democracy—a system where every vote counts, but some states matter just a bit more than others.

@@

Recency weights are calculated based on the number of days elapsed since the poll ended, utilizing an exponential decay function to give greater weight to more recent polls. Additionally, a sample size weight is implemented, capping the weights at a maximum of 2,300 responses to reflect the reliability of larger sample sizes.

@@

## Model validation

```{r}
# Set seed for reproducibility
set.seed(123)

# Sample indices for training and testing sets
train_index <- sample(seq_len(nrow(data)), size = 0.8 * nrow(data))

# Split the data into training and testing sets
train_data <- data[train_index, ]
test_data <- data[-train_index, ]

# Fit the weighted linear regression model
first_model_test <- lm(pct ~ national_poll + pollster + population, 
                     data = data, 
                  data = train_data)

# Predict on test data
predictions <- predict(first_model_test, newdata = test_data)

# Calculate RMSE using the correct outcome variable
# Assuming 'pct' is the actual observed outcome variable in test_data
rmse <- sqrt(mean((predictions - test_data$pct)^2, na.rm = TRUE))

# Print RMSE
print(paste("RMSE:", rmse))

@@

@fig-stanareyouokay-1 is a residual plot of un-weighted model. It shows... This suggests...

@fig-stanareyouokay-2 is a Q-Q plot of un-weighted model plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
#residual plot
plot(first_model, which = 1,main ="Residual Plot of unweighted linear regression model residualsmodel" )
# Q-Q plot for the residuals of the model
qqnorm(residuals(first_model),main= "Q-Q Plot of unweighted linear regression model residuals
")
qqline(residuals(first_model), col = "red")

@@
